---
title: "Cockroach Dashboard"
output: github_document
---

```{r, message = FALSE, warning = FALSE, include=FALSE, echo=FALSE}
library(flexdashboard)
library(tidyverse)
library(viridis)
library(leaflet)
library(plotly)
library(shiny)
library(sf)
library(sp)
library(rgdal)
library(rsconnect)
```

Below is the same code for cleaning data as used in our `p8105_final` repository. Added is the code to save `tidy_asthma_sf_ll` as an `.Rda` file so it can be easily accessed for the shiny app.

The `all_available_data.csv` can be downloaded (here)[http://a816-dohbesp.nyc.gov/IndicatorPublic/EPHTCsv/AllAvailableData.zip]. The website with all datasets is (here)[http://a816-dohbesp.nyc.gov/IndicatorPublic/BuildATable.aspx].

I used a different shapefiles source than the nyc.gov files because the Env & Health Data Portal file was inconsistent - combined certain neighborhoods that weren't combined in the rest of the datasets. I got the [shape files](https://www1.nyc.gov/site/doh/data/health-tools/maps-gis-data-files-for-download.page) from a different nyc.gov page.

```{r, message = FALSE, warning = FALSE}
all_data = 
read.csv(file = "./roach_shiny/data/all_available_data.csv") %>% 
  janitor::clean_names() %>% 
  filter(name %in% c("Homes with Cockroaches", 
                     "Poverty", 
                     "Public School Children (5-14 Yrs Old) with Asthma", 
                     "Public School Children (5-14 Yrs Old) with Persistent Asthma", 
                     "Asthma Hospitalizations (Children 5 to 17 Yrs Old)"),
        geo_type_name %in% c("UHF42", "Borough"), 
        measure %in% c("Percent", "Rate")
         ) %>% 
  mutate(geo_join_id = as.character(geo_join_id),
         geo_join_id = if_else(geo_place_name == "Bensonhurst - Bay Ridge", "209", geo_join_id),
         geo_join_id = as.factor(geo_join_id)) %>% 
  pivot_wider(
    names_from = name, 
    values_from = data_value
    ) %>%
  mutate(time_period = recode(time_period, "2010-2011" = "2011", "2011-2012" = "2012", "2012-2013" = "2013", "2013-2014" = "2014", "2010-14" = "2011")) %>%
  filter(time_period %in% c("2011", "2012", "2013", "2014")) %>% 
  janitor::clean_names() %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma, asthma_hospitalizations_children_5_to_17_yrs_old)

df1 = all_data  %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, homes_with_cockroaches) %>% 
  drop_na()

df2 = all_data %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, poverty) %>% 
  drop_na()

df3 = all_data %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, public_school_children_5_14_yrs_old_with_asthma) %>% 
  drop_na()

df4 = all_data %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, public_school_children_5_14_yrs_old_with_persistent_asthma) %>% 
  drop_na()

df5 = all_data %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, measure, indicator_id, asthma_hospitalizations_children_5_to_17_yrs_old) %>% 
  drop_na()

## Joining
tidy_asthma = 
  full_join(df1, df2, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  full_join(df3, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  full_join(df4, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  full_join(df5, by = c("geo_type_name", "geo_join_id", "time_period", "geo_place_name")) %>% 
  select(geo_type_name, geo_join_id, time_period, geo_place_name, homes_with_cockroaches, poverty, public_school_children_5_14_yrs_old_with_asthma, public_school_children_5_14_yrs_old_with_persistent_asthma, asthma_hospitalizations_children_5_to_17_yrs_old) %>% 
  view()

#importing shapefiles
shape = st_read("./roach_shiny/data/shapefiles/", quiet = TRUE)

#merging with the tidy dataset
tidy_asthma_sf = merge(shape, tidy_asthma, by.x = "UHFCODE", by.y = "geo_join_id") %>% 
  janitor::clean_names() %>% 
  mutate(uhfcode = as.character(uhfcode),
         time_period = as.character(time_period),
         homes_with_cockroaches = as.numeric(homes_with_cockroaches))

#changing geometry to latitude and longitude
tidy_asthma_sf_ll = st_transform(tidy_asthma_sf, "+proj=longlat +datum=WGS84")

#saving as an R dataframe so the shiny app can access it
save(tidy_asthma_sf_ll, file = "tidy_asthma_sf_ll.Rda")
```
